---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load packages:
```{r}
source("scripts/pkgs.R")
```

# load data = the results of predicted TF's binding sites in 1500 bp promoters in Arabidopsis
```{r}
TF_scan <- read_csv("analysed_data/TF_rc_scan_results_promoters1500_TF_family.csv")
```

# Update "start" and "stop" coordinates for "-" strand to be always start < stop
```{r}
TF_scan_updated <- 
TF_scan %>% 
  mutate(start_updated = if_else(strand == "-", stop, start), stop_updated = if_else(strand == "-", start, stop)) 

# Create list of all TF; Consider to sort next time
TF_list <- sort(unique(TF_scan$motif))

# Calculate total number of binding site for each TF; order as TF_list 
N_TF_bs <- TF_scan %>% 
  count(motif)
```

# In order to find overlaps in TF binding sites use GRanges
# Create GRanges obgect; use ID number instead of "chromosome names" as seqnames to use later to find overlaps
```{r}
Ara_gr <- GRanges(seqnames = Rle(TF_scan_updated$sequence),
              ranges = IRanges(TF_scan_updated$start_updated, end = TF_scan_updated$stop_updated),
               strand = Rle(strand(TF_scan_updated$strand)), select(TF_scan_updated, -starts_with("stop"), -starts_with("start"), -strand ))
Ara_gr
```


# Quick test of consept for finding "overlaps"
```{r}
Ara_gr_TF1 <- 
  Ara_gr[mcols(Ara_gr)$motif == TF_list[1]]

Ara_gr_TF2 <- 
  Ara_gr[mcols(Ara_gr)$motif == TF_list[2]]

findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")
queryHits(findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")) #TF1
subjectHits(findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")) #TF2
countOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")
sum(countOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any"))

test_1 <- findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")
nLnode(test_1)
nRnode(test_1)

```


```{r}

```



```{r}
Overlap_TF_bs <- function(n,m){
  TF1 <- TF_list[n]
  TF2 <- TF_list[m]
  
  Ara_gr_TF1 <- 
  Ara_gr[mcols(Ara_gr)$motif == TF1]

Ara_gr_TF2 <- 
  Ara_gr[mcols(Ara_gr)$motif == TF2]

Overlap_TF <- findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")

message(paste(n,m))

return(
data.frame("motif_1" = TF1, "motif_2" = TF2, 
           "overlap" = length(queryHits(Overlap_TF)), 
            "overlap_TF_1" = length(unique(queryHits(Overlap_TF))),
              "overlap_TF_2" = length(unique(subjectHits(Overlap_TF))),
                "tot_TF1" = nLnode(Overlap_TF),
                 "tot_TF2" = nRnode(Overlap_TF), 
                   stringsAsFactors = FALSE)
      )
}

Overlap_TF_bs(8,9)
###########

n = length(TF_list)
#n = 50
x <- Sys.time()

Results <- lapply(seq_along(TF_list[1:(n-1)]), function(y)(
  lapply(seq(y+1,n), function(x)Overlap_TF_bs(n=y, m=x))
)

)


Sys.time()-x 

Results_df <- do.call(bind_rows, Results)
Results_df <- Results_df %>% 
  mutate( pct_overlap_TF1 = round(overlap_TF_1/tot_TF1,2 ),
                   pct_overlap_TF2 = round(overlap_TF_2/tot_TF2, 2)) %>% 
 arrange(desc(pct_overlap_TF1, pct_overlap_TF2))


```


```{r warning=FALSE, include=FALSE}

n = length(TF_list)

x <- Sys.time()
# last run 600:(n-1)

Results_1_300 <- lapply(seq_along(TF_list[1:300]), function(y)(
  lapply(seq(y+1,n), function(x)Overlap_TF_bs(n=y, m=x))
)

)


Sys.time()-x 

Results_df <- do.call(bind_rows, Results_1_300)
Results_df <- Results_df %>% 
  mutate( pct_overlap_TF1 = round(overlap_TF_1/tot_TF1,2 ),
                   pct_overlap_TF2 = round(overlap_TF_2/tot_TF2, 2)) %>% 
 arrange(desc(pct_overlap_TF1, pct_overlap_TF2))


write_csv(Results_df, "output_data/Results_1_300.csv")

```


```{r}
saveRDS(file = here(paste0("output_data","Results_1_300",".rda")) , Results_1_300)
```

```{r}
# read_rds("output_data/output_dataResults_701.rda") #Stack computer
```



