---
title: "Analysis of TF binding site similarity"
output: html_document
editor_options: 
  chunk_output_type: inline
---

Aim: To compare predicted binding sites of TFs and find % of binding sites overlap by 4 or more base pairs
data_input: analysed_data/TF_rc_scan_results_promoters1500_TF_family.csv
            Brapa_3/analysed_data/TF_rc_scan_results_Brapa3_promoters1500_TF_family.csv
            

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE)
```

# load packages:
```{r}
source("scripts/pkgs.R")
```

# load data = the results of predicted TF's binding sites in 1500 bp promoters in Arabidopsis
```{r}
TF_scan <- read_csv("../TF_binding/analysed_data/TF_rc_scan_results_promoters1500_TF_family.csv")
```

# Update "start" and "stop" coordinates for "-" strand to be always start < stop
```{r}
TF_scan_updated <- 
TF_scan %>% 
  mutate(start_updated = if_else(strand == "-", stop, start), stop_updated = if_else(strand == "-", start, stop)) 

# Create list of all TF; Consider to sort next time
TF_list <- sort(unique(TF_scan$motif))

# Calculate total number of binding site for each TF; order as TF_list 
N_TF_bs <- TF_scan %>% 
  count(motif)
```


# In order to find overlaps in TF binding sites use GRanges
# Create GRanges obgect; use ID number instead of "chromosome names" as seqnames to use later to find overlaps
```{r}
Ara_gr <- GRanges(seqnames = Rle(TF_scan_updated$sequence),
              ranges = IRanges(TF_scan_updated$start_updated, end = TF_scan_updated$stop_updated),
               strand = Rle(strand(TF_scan_updated$strand)), select(TF_scan_updated, -starts_with("stop"), -starts_with("start"), -strand ))
Ara_gr
```


##------*Development*--------##
##------ Quick test of consept for finding "overlaps"
```{r}
Ara_gr_TF1 <- 
  Ara_gr[mcols(Ara_gr)$motif == TF_list[1]]

Ara_gr_TF2 <- 
  Ara_gr[mcols(Ara_gr)$motif == TF_list[2]]

findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")
queryHits(findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")) #TF1
subjectHits(findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")) #TF2
countOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")
sum(countOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any"))

test_1 <- findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")
nLnode(test_1)
nRnode(test_1)

```


##------ Development of function to include more info, like number of TF binding sites having overlap (not the same as tot N of overlaps)
```{r}
Ara_gr_TF1 <- 
  Ara_gr[mcols(Ara_gr)$motif == "ABI3VP1_tnt.VRN1_col_a_m1"]

Ara_gr_TF2 <- 
  Ara_gr[mcols(Ara_gr)$motif == "ABI3VP1_tnt.VRN1_colamp_a_m1"]

findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")
queryHits(findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")) #TF1
subjectHits(findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")) #TF2
countOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")
sum(countOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any"))

test_1 <- findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")
nLnode(test_1)
nRnode(test_1)

findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")[1:100]

Ara_gr_TF1[1]
Ara_gr_TF2[54:58]

findOverlaps(Ara_gr_TF2, Ara_gr_TF1, minoverlap=4L, type = "any")


length(unique(queryHits(findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any"))))
length(queryHits(findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")))

length(unique(subjectHits(findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any"))))
length(subjectHits(findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")))

which(TF_list=="ABI3VP1_tnt.VRN1_col_a_m1")
which(TF_list=="ABI3VP1_tnt.VRN1_colamp_a_m1")
```

##------

#--Write function *Overlap_TF_bs* to calculate number of binding sites overlapping in two TFs (pairwised). The output will include more info, like the number of TF binding sites having overlap (not the same as tot N of overlaps), number of overlaps
```{r}

Overlap_TF_bs <- function(n,m){
  TF1 <- TF_list[n]
  TF2 <- TF_list[m]
  
  Ara_gr_TF1 <- 
  Ara_gr[mcols(Ara_gr)$motif == TF1]

Ara_gr_TF2 <- 
  Ara_gr[mcols(Ara_gr)$motif == TF2]

Overlap_TF <- findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")

message(paste(n,m))

return(
data.frame("motif_1" = TF1, "motif_2" = TF2, 
           "overlap" = length(queryHits(Overlap_TF)), 
            "overlap_TF_1" = length(unique(queryHits(Overlap_TF))),
              "overlap_TF_2" = length(unique(subjectHits(Overlap_TF))),
                "tot_TF1" = nLnode(Overlap_TF),
                 "tot_TF2" = nRnode(Overlap_TF), 
                   stringsAsFactors = FALSE)
      )
}

Overlap_TF_bs(8,9)

```



###-- Run *Overlap_TF_bs* on parts of the TF_list

```{r}

n = length(TF_list)

x <- Sys.time()
# last run 600:(n-1)

Results_1_300 <- lapply(seq_along(TF_list[1:300]), function(y)(
  lapply(seq(y+1,n), function(x)Overlap_TF_bs(n=y, m=x))
)

)


Sys.time()-x 

Results_df <- do.call(bind_rows, Results_1_300)

```


```{r}
write_csv(Results_df, "output_data/Results_1_300.csv")
saveRDS(file = here(paste0("output_data","Results_1_300",".rda")) , Results_1_300)
```

```{r}
# read_rds("output_data/output_dataResults_701.rda") #Stack computer
```

```{r}
###########

n = length(TF_list)

x <- Sys.time()
# last run 601:(n-1)

Results_301_600 <- lapply(seq(301,600), function(y)(
  lapply(seq(y+1,n), function(x)Overlap_TF_bs(n=y, m=x))
)

)


Sys.time()-x 

############
n = length(TF_list)

x <- Sys.time()
# last run 601:(n-1)

Results_601_700 <- lapply(seq(601,700), function(y)(
  lapply(seq(y+1,n), function(x)Overlap_TF_bs(n=y, m=x))
)

)


Sys.time()-x 

############
n = length(TF_list)

x <- Sys.time()
# last run 601:(n-1)

Results_701_871 <- lapply(seq(701,871), function(y)(
  lapply(seq(y+1,n), function(x)Overlap_TF_bs(n=y, m=x))
)

)


Sys.time()-x 

##########
```

##-- If required to run *Overlap_TF_bs* on wholl data set at once---#########
```{r eval=FALSE, include=FALSE}
n = length(TF_list)
x <- Sys.time()

Results_all <- lapply(seq_along(TF_list[1:(n-1)]), function(y)(
  unlist(lapply(seq(y+1,n), function(x)test(n=y, m=x)))
)

)

Sys.time()-x 
```



## Combine all the partial runs (have been done in diiferent projects in parallel)
## Calculate the percentage of overlaping binding sites to all TF pairedwise, on all data at the end
```{r}
Results_df_2 <- do.call(bind_rows, c(Results_301_600, Results_601_700, Results_701_871))
# Results_df_2 <- read_csv("../TF_binding/output_data/Results_301_782.csv")


```

###-- Save output as ".rda" and ".csv"
```{r}
saveRDS(file = here(paste0("output_data","Results_301_600",".rda")) , Results_301_600)
saveRDS(file = here(paste0("output_data","Results_061_700",".rda")) , Results_601_700)
saveRDS(file = here(paste0("output_data","Results_701",".rda")) , Results_701_871)
write_csv(Results_df_2, "../TF_binding/output_data/Results_301_782.csv")

```


##--read saved output
```{r}
Results_1_300 <- read_csv(here("output_data", "Results_1_300.csv"))
Results_301_782<- read_csv(here("output_data", "Results_301_782.csv"))



Results_all_df <- bind_rows(Results_1_300, Results_301_782) %>%
  mutate( pct_overlap_TF1 = round(overlap_TF_1/tot_TF1,2 ),
                   pct_overlap_TF2 = round(overlap_TF_2/tot_TF2, 2)) 

Results_all_df %>%
 arrange(desc(pct_overlap_TF1, pct_overlap_TF2))

# write_csv(Results_all_df, "../TF_binding/output_data/Results_overlap_TFs_all.csv")
```


##------*Development*--------##
###-- Write function calculating number of overlaps between binding sites of certain TF and all the rest of TF from the list "TF_list"
###-*Note This functoin not suitable for calculation of %  of overlaped TF bs between two TFs*
###-Output total namber of overlapes (if TF1 has n overlaps with TF2, it will count it as n)
```{r}

Overlap_TF_bs <- function(n,m){
  Ara_gr_TF1 <- 
  Ara_gr[mcols(Ara_gr)$motif == TF_list[n]]

Ara_gr_TF2 <- 
  Ara_gr[mcols(Ara_gr)$motif == TF_list[m]]

message(paste(n,m))

return(
data.frame("motif_1" = TF_list[n], "motif_2" = TF_list[m], "overlap" = sum(countOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")), stringsAsFactors = FALSE)

)

}

# --- Run the following for pairwise comparesent af all TFs ##########

n = length(TF_list)
x <- Sys.time()

Results <- lapply(seq_along(TF_list[1:(n-1)]), function(y)(
  lapply(seq(y+1,n), function(x)Overlap_TF_bs(n=y, m=x))
)

)


Sys.time()-x

Results_df <- do.call(bind_rows, Results)

```


##--Test that all TFs have been compared properly
```{r cache=TRUE, include=FALSE, results='hide'}
Overlap_TF_bs_test <- function(n,m){
  TF1 <- TF_list[n]
  TF2 <- TF_list[m]
  
#   Ara_gr_TF1 <- 
#   Ara_gr[mcols(Ara_gr)$motif == TF1]
# 
# Ara_gr_TF2 <- 
#   Ara_gr[mcols(Ara_gr)$motif == TF2]

# Overlap_TF <- findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")

message(paste(n,m))

return(
data.frame("motif_1" = TF1, "motif_2" = TF2,
           "n_m" = paste(n,m), stringsAsFactors = FALSE)
      )
}

#########
Overlap_TF_bs_test(8,9)

##########
n = length(TF_list)
x <- Sys.time()

Results_all_test <- lapply(seq_along(TF_list[1:(n-1)]), function(y)(
  lapply(seq(y+1,n), function(x)Overlap_TF_bs_test(n=y, m=x))
)

)

Sys.time()-x 

Results_all_test_df <- do.call(bind_rows, Results_all_test)
all(Results_all_df$motif_2 == Results_all_test_df$motif_2) #should be TRUE
all(Results_all_df$motif_1 == Results_all_test_df$motif_1)
```

Time difference of 3.931119 mins
### Consider to parallelirise:

library(parallel)
library(MASS)

mclapply(starts, fx, mc.cores = numCores)
