---
title: "Analysis of TF binding site similarity in Arabidopsis"
output: html_notebook
---


Aim: To compare predicted binding sites of TFs and find % of binding sites overlap by 4 or more base pairs
data_input: Arabidopsis: analysed_data/TF_rc_scan_results_promoters1500_TF_family.csv 
            Brassica: Brapa_3/analysed_data/TF_rc_scan_results_Brapa3_promoters1500_TF_family.csv
   

# load packages:
```{r}
source("scripts/pkgs.R")
```

# load data = the results of predicted TF's binding sites in 1500 bp promoters in Arabidopsis
```{r}
TF_scan <- read_csv("../TF_binding/analysed_data/TF_rc_scan_results_promoters1500_TF_family.csv")
```

# Update "start" and "stop" coordinates for "-" strand to be always start < stop
```{r}
TF_scan_updated <- 
TF_scan %>% 
  mutate(start_updated = if_else(strand == "-", stop, start), stop_updated = if_else(strand == "-", start, stop)) 

# Create list of all TF; Consider to sort next time
TF_list <- sort(unique(TF_scan$motif))

# Calculate total number of binding site for each TF; order as TF_list 
N_TF_bs <- TF_scan %>% 
  count(motif)
```


# In order to find overlaps in TF binding sites use GRanges
# Create GRanges obgect; use ID number instead of "chromosome names" as seqnames to use later to find overlaps
```{r}
Ara_gr <- GRanges(seqnames = Rle(TF_scan_updated$sequence),
              ranges = IRanges(TF_scan_updated$start_updated, end = TF_scan_updated$stop_updated),
               strand = Rle(strand(TF_scan_updated$strand)), select(TF_scan_updated, -starts_with("stop"), -starts_with("start"), -strand ))
Ara_gr
```


##------*Development*--------##
##------ Quick test of consept for finding "overlaps"
```{r}
Ara_gr_TF1 <- 
  Ara_gr[mcols(Ara_gr)$motif == TF_list[1]]

Ara_gr_TF2 <- 
  Ara_gr[mcols(Ara_gr)$motif == TF_list[2]]

findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")
queryHits(findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")) #TF1
subjectHits(findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")) #TF2
countOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")
sum(countOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any"))

test_1 <- findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")
nLnode(test_1)
nRnode(test_1)

```


##------ Development of function to include more info, like number of TF binding sites having overlap (not the same as tot N of overlaps)
```{r}
Ara_gr_TF1 <- 
  Ara_gr[mcols(Ara_gr)$motif == "ABI3VP1_tnt.VRN1_col_a_m1"]

Ara_gr_TF2 <- 
  Ara_gr[mcols(Ara_gr)$motif == "ABI3VP1_tnt.VRN1_colamp_a_m1"]

findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")
queryHits(findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")) #TF1
subjectHits(findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")) #TF2
countOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")
sum(countOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any"))

test_1 <- findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")
nLnode(test_1)
nRnode(test_1)

findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")[1:100]

Ara_gr_TF1[1]
Ara_gr_TF2[54:58]

findOverlaps(Ara_gr_TF2, Ara_gr_TF1, minoverlap=4L, type = "any")


length(unique(queryHits(findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any"))))
length(queryHits(findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")))

length(unique(subjectHits(findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any"))))
length(subjectHits(findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")))

which(TF_list=="ABI3VP1_tnt.VRN1_col_a_m1")
which(TF_list=="ABI3VP1_tnt.VRN1_colamp_a_m1")
```

##------

#--Write function *Overlap_TF_bs* to calculate number of binding sites overlapping in two TFs (pairwised). The output will include more info, like the number of TF binding sites having overlap (not the same as tot N of overlaps), number of overlaps
```{r}

Overlap_TF_bs <- function(n,m){
  TF1 <- TF_list[n]
  TF2 <- TF_list[m]
  
  Ara_gr_TF1 <- 
  Ara_gr[mcols(Ara_gr)$motif == TF1]

Ara_gr_TF2 <- 
  Ara_gr[mcols(Ara_gr)$motif == TF2]

Overlap_TF <- findOverlaps(query = Ara_gr_TF1, subject = Ara_gr_TF2, minoverlap = 4L, type = "any")

message(paste(n,m))

return(
data.frame("motif_1" = TF1, "motif_2" = TF2, 
           "overlap" = length(queryHits(Overlap_TF)), 
            "overlap_TF_1" = length(unique(queryHits(Overlap_TF))),
              "overlap_TF_2" = length(unique(subjectHits(Overlap_TF))),
                "tot_TF1" = nLnode(Overlap_TF),
                 "tot_TF2" = nRnode(Overlap_TF), 
                   stringsAsFactors = FALSE)
      )
}

Overlap_TF_bs(8,9)

```



###-- Run *Overlap_TF_bs* on parts of the TF_list

```{r}

n = length(TF_list)

x <- Sys.time()
# last run 600:(n-1)

Results_1_300 <- lapply(seq_along(TF_list[1:300]), function(y)(
  lapply(seq(y+1,n), function(x)Overlap_TF_bs(n=y, m=x))
)

)


Sys.time()-x 

Results_df <- do.call(bind_rows, Results_1_300)

```


```{r}
write_csv(Results_df, "output_data/Results_1_300.csv")
saveRDS(file = here(paste0("output_data","Results_1_300",".rda")) , Results_1_300)
```

```{r}
# read_rds("output_data/output_dataResults_701.rda") #Stack computer
```

```{r}
###########

n = length(TF_list)

x <- Sys.time()
# last run 601:(n-1)

Results_301_600 <- lapply(seq(301,600), function(y)(
  lapply(seq(y+1,n), function(x)Overlap_TF_bs(n=y, m=x))
)

)


Sys.time()-x 

############
n = length(TF_list)

x <- Sys.time()
# last run 601:(n-1)

Results_601_700 <- lapply(seq(601,700), function(y)(
  lapply(seq(y+1,n), function(x)Overlap_TF_bs(n=y, m=x))
)

)


Sys.time()-x 

############
n = length(TF_list)

x <- Sys.time()
# last run 601:(n-1)

Results_701_871 <- lapply(seq(701,871), function(y)(
  lapply(seq(y+1,n), function(x)Overlap_TF_bs(n=y, m=x))
)

)


Sys.time()-x 

##########
```

##-- If required to run *Overlap_TF_bs* on wholl data set at once---#########
```{r eval=FALSE, include=FALSE}
n = length(TF_list)
x <- Sys.time()

Results_all <- lapply(seq_along(TF_list[1:(n-1)]), function(y)(
  unlist(lapply(seq(y+1,n), function(x)test(n=y, m=x)))
)

)

Sys.time()-x 
```



## Combine all the partial runs (have been done in diiferent projects in parallel)
## Calculate the percentage of overlaping binding sites to all TF pairedwise, on all data at the end
```{r}
Results_df_2 <- do.call(bind_rows, c(Results_301_600, Results_601_700, Results_701_871))
# Results_df_2 <- read_csv("../TF_binding/output_data/Results_301_782.csv")


```

###-- Save output as ".rda" and ".csv"
```{r}
saveRDS(file = here(paste0("output_data","Results_301_600",".rda")) , Results_301_600)
saveRDS(file = here(paste0("output_data","Results_061_700",".rda")) , Results_601_700)
saveRDS(file = here(paste0("output_data","Results_701",".rda")) , Results_701_871)
write_csv(Results_df_2, "../TF_binding/output_data/Results_301_782.csv")

```


##--read saved output
```{r}
Results_1_300 <- read_csv(here("output_data/02_TF_bs_similarity_dimension_reduction_Arabidopsis/", "Results_1_300.csv"))
Results_301_782<- read_csv(here("output_data/02_TF_bs_similarity_dimension_reduction_Arabidopsis/", "Results_301_782.csv"))



Results_all_df <- bind_rows(Results_1_300, Results_301_782) %>%
  mutate( pct_overlap_TF1 = round(overlap_TF_1/tot_TF1,2 ),
                   pct_overlap_TF2 = round(overlap_TF_2/tot_TF2, 2)) 

Results_all_df %>%
 arrange(desc(pct_overlap_TF1, pct_overlap_TF2))

# write_csv(Results_all_df, "../TF_binding/output_data/Results_overlap_TFs_all.csv")
```


##------*Development*--------##
###-- Write function calculating number of overlaps between binding sites of certain TF and all the rest of TF from the list "TF_list"
###-*Note This functoin not suitable for calculation of %  of overlaped TF bs between two TFs*
###-Output total namber of overlapes (if TF1 has n overlaps with TF2, it will count it as n)
```{r}

Overlap_TF_bs <- function(n,m){
  Ara_gr_TF1 <- 
  Ara_gr[mcols(Ara_gr)$motif == TF_list[n]]

Ara_gr_TF2 <- 
  Ara_gr[mcols(Ara_gr)$motif == TF_list[m]]

message(paste(n,m))

return(
data.frame("motif_1" = TF_list[n], "motif_2" = TF_list[m], "overlap" = sum(countOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")), stringsAsFactors = FALSE)

)

}

# --- Run the following for pairwise comparesent af all TFs ##########

n = length(TF_list)
x <- Sys.time()

Results <- lapply(seq_along(TF_list[1:(n-1)]), function(y)(
  lapply(seq(y+1,n), function(x)Overlap_TF_bs(n=y, m=x))
)

)


Sys.time()-x

Results_df <- do.call(bind_rows, Results)

```


##--Test that all TFs have been compared properly
```{r cache=TRUE, include=FALSE, results='hide'}
Overlap_TF_bs_test <- function(n,m){
  TF1 <- TF_list[n]
  TF2 <- TF_list[m]
  
#   Ara_gr_TF1 <- 
#   Ara_gr[mcols(Ara_gr)$motif == TF1]
# 
# Ara_gr_TF2 <- 
#   Ara_gr[mcols(Ara_gr)$motif == TF2]

# Overlap_TF <- findOverlaps(Ara_gr_TF1, Ara_gr_TF2, minoverlap=4L, type = "any")

message(paste(n,m))

return(
data.frame("motif_1" = TF1, "motif_2" = TF2,
           "n_m" = paste(n,m), stringsAsFactors = FALSE)
      )
}

#########
Overlap_TF_bs_test(8,9)

##########
n = length(TF_list)
x <- Sys.time()

Results_all_test <- lapply(seq_along(TF_list[1:(n-1)]), function(y)(
  lapply(seq(y+1,n), function(x)Overlap_TF_bs_test(n=y, m=x))
)

)

Sys.time()-x 

Results_all_test_df <- do.call(bind_rows, Results_all_test)
all(Results_all_df$motif_2 == Results_all_test_df$motif_2) #should be TRUE
all(Results_all_df$motif_1 == Results_all_test_df$motif_1)
```

Time difference of 3.931119 mins
### Consider to parallelirise:

library(parallel)
library(MASS)

mclapply(starts, fx, mc.cores = numCores)



#####---*TEST*------########

##--read saved output



###--Check if tot N bs is OK
####--compare calculated by "Overlap_TF_bs" and  "N_TF_bs"



```{r}
Results_all_test <- read_csv("../TF_binding/output_data/02_TF_bs_similarity_dimension_reduction_Arabidopsis/Results_overlap_TFs_all.csv")


test <- bind_rows(
# part_1
 (Results_all_test %>% 
   select(motif_1, motif_2, overlap = overlap_TF_1, tot = tot_TF1)),
   
# part_2 
(Results_all_test %>% 
   select(motif_2 = motif_1, motif_1 = motif_2, overlap = overlap_TF_2, tot = tot_TF2)),
# part_3
 (N_TF_bs %>% 
  select(motif_1 = motif, tot = n) %>% 
  mutate(motif_2 = motif_1, overlap = tot))
) %>%

  mutate(pct_overlap = round(overlap/tot*100,3)) %>%
  select(-overlap, -tot) %>%
  spread(motif_2, pct_overlap)




which(is.na(test), arr.ind = TRUE)
test[1,2]

```

```{r}
source("~/Work/2019/Analysis/Utilities_R/Functions/My_functions.R")

data_matrix <- test
Title_plot <- "hclust_test" # main= Title for the plot; string
Path_to_output_results <- "../TF_binding/output_data/"
Path_to_plot <- "../TF_binding/plots/"

```


#### Clustering TF binding site by hclust.person / columns
```{r eval=FALSE}
#arguments:
hclust_data <- data_matrix %>% column_to_rownames(.,var = "motif_1") #Input matrix to cluster

k_row = 20  # n of clusters in cutree 
k_col = 20


my_brewer_par <- c(brewer.pal(12,'Paired'),brewer.pal(8,'Set1'))


hclust_rows <- hclust.person.rows(hclust_data, k_row) #from my_functions
names(hclust_rows)
#[1] "Rowv_arg" "RowSideColors_arg" "Cluster_rows"   

hclust_columns <- hclust.person.columns(hclust_data, k_col)
names(hclust_columns)
#[1] "Colv_arg" "ColSideColors_arg" "Cluster_columns"  

names(hclust_data)


```


##-- Create heatmap
###-- Define arguments
```{r eval=FALSE}
#arguments:
data_plot <- hclust_data #Input matrix to plot

column_names <- str_remove(colnames(data_plot), "\\..*$") #labCol=column_names; vector of the names to be ploted for each column in heatmap

#only for dendograms add the folowing arguments to create sep horizontal lines in heatmap; #line/break after columns/row number in vector
## rows

row_sep <- table(hclust_rows$Cluster_rows)

## columns
ind_c <- match(labels(hclust_columns$Colv_arg), colnames(hclust_data))
col_sep <- table(hclust_columns$Cluster_columns)[match(unique(hclust_columns$Cluster_columns[ind_c]),
           names(table(hclust_columns$Cluster_columns)))]

rowsep_arg <- sep_calc(row_sep)
colsep_arg <- sep_calc(col_sep)


########
length(hclust_rows$RowSideColors_arg)
dim(data_plot)
names(data_plot)

table(hclust_rows$RowSideColors_arg)




```


##-- Plot heatmap dendogram rows and columns 
```{r eval=FALSE}
# pdf(paste0(Path_to_plot, Title_plot,"_dend_TF", ".pdf"))
# par(cex.main=1) #controls the size of the title
#data_heatmap <- 
                heatmap.2(as.matrix(data_plot),
                          Rowv = hclust_rows$Rowv_arg,
                          #Rowv = FALSE,
                          Colv = hclust_columns$Colv_arg,
                          dendrogram = 'column',
                          #dendrogram = 'both',
                          scale = "column",
                          col = colorRampPalette(c("blue","white", "red"))(n = 100),
                          breaks=seq(-2, 2,length.out=101), 
                          trace = 'none',         
                          labRow=NA,  
                          labCol=NA,
                          main=Title_plot,
                          RowSideColors = hclust_rows$RowSideColors_arg, 
                          ColSideColors = hclust_columns$ColSideColors_arg, 
                          key=TRUE,
                          density.info = "none",  #remove the hist from key
                          lhei = c(2, 8), # this makes the colour-key legend a little thinner,
                          sepwidth=c(0.05,0.05),
                          sepcolor="white",
                          # colsep = colsep_arg,
                          # rowsep = rowsep_arg,
                          # the margins command sets the width of the white space around the plot. 
                          #The first element is the bottom margin and the second is the right margin
                          margins = c(12, 2)
)
#dev.off()



heatmap.2(as.matrix(data_plot),
                          Rowv = hclust_rows$Rowv_arg,
                          Colv = hclust_columns$Colv_arg,
                          dendrogram = 'both',
                          #scale = "column",
                          col = colorRampPalette(c("blue","white", "red"))(n = 100),
                          breaks=seq(0, 100,length.out=101), 
                          trace = 'none',         
                          labRow=NA,  
                          labCol=NA,
                          main=Title_plot,
                          RowSideColors = hclust_rows$RowSideColors_arg, 
                          ColSideColors = hclust_columns$ColSideColors_arg, 
                          key=TRUE,
                          density.info = "none",  #remove the hist from key
                          #lhei = c(2, 8), # this makes the colour-key legend a little thinner,
                          sepwidth=c(0.05,0.05),
                          sepcolor="white",
                          # colsep = colsep_arg,
                          # rowsep = rowsep_arg,
                          # the margins command sets the width of the white space around the plot. 
                          #The first element is the bottom margin and the second is the right margin
                         # margins = c(12, 2)
)


```


```{r}

data_matrix$cluster <- hclust_rows$Cluster_rows[1:10]
data_plot_reordered <-  bind_rows(data_matrix, hclust_columns$Cluster_columns)

length(hclust_rows$Cluster_rows)
all(row.names(hclust_data)==names(hclust_rows$Cluster_rows))


hclust_data_claster_reordered_v1 <- 
  data_matrix %>% 
  mutate(cluster = hclust_rows$Cluster_rows) %>% 
   bind_rows(. , hclust_columns$Cluster_columns)


# %>% 
#   select(order.dendrogram(hclust_rows$Rowv_arg), cluster)

hclust_data_claster_reordered[,870:873]
hclust_data[,870:872]

hclust_data_claster_reordered_v1[870:873,]
hclust_data_claster_reordered_v1[873,1] <- "cluster_col"

hclust_data_claster_reordered_v1[,c(1,874)]
hclust_data_claster_reordered_v1[order.dendrogram(hclust_rows$Rowv_arg), order.dendrogram(hclust_columns$Colv_arg)]
length(order.dendrogram(hclust_rows$Rowv_arg))
length(order.dendrogram(hclust_columns$Colv_arg))

```




```{r}
Cluster_rows <- as.data.frame(list(hclust_rows$Cluster_rows), col.names = c("cluster_rows")) %>% rownames_to_column(., var = "motif_1")
Cluster_col <- as.data.frame(list(hclust_columns$Cluster_columns), col.names = c("cluster_col")) %>% rownames_to_column(., var = "motif_2")

hclust_data_claster_reordered <- 
  data_matrix %>% 
  gather( "motif_2", "pct", -motif_1) %>% 
  full_join(., Cluster_rows, by = "motif_1") %>% 
  full_join(., Cluster_col, by = "motif_2")

test <- hclust_data_claster_reordered %>% # filter(cluster_rows == 1) %>% 
  group_by(cluster_rows, cluster_col) %>% 
  summarise(mean = mean(pct)) 

hclust_data_claster_reordered %>% 
  filter(cluster_rows == 20, cluster_col == 20) %>% arrange(pct)

```




```{r}
kmeans_rows <- kmeans(hclust_data, 40)
kmeans_cols <- kmeans(t(hclust_data), 40)

# names(sort(kmeans_rows$cluster[1:100])) 

# base::match(names(order(kmeans_rows$cluster)),colnames(hclust_data) )
# kmeans_cols$cluster

# hclust_data[kmeans_rows$cluster, kmeans_cols$cluster]
#test_2[order.dendrogram(hclust_rows$Rowv_arg), order.dendrogram(hclust_columns$Colv_arg)]

heatmap(as.matrix(hclust_data[order(kmeans_rows$cluster), order(kmeans_cols$cluster)]),
                          Rowv = kmeans_rows$cluster,
                          Colv = kmeans_cols$cluster,
                          #scale = "column",
                          col = colorRampPalette(c("blue","white", "red"))(n = 100),
                          breaks=seq(-3,3,length.out=101),      
                          labRow=NA,  
                          labCol=NA,
                          main=Title_plot,
                          #RowSideColors = brewer_par[clust_data$cluster], 
                          #ColSideColors = hclust_columns$ColSideColors_arg
)
```

```{r}
hclust_data[kmeans_rows$cluster, kmeans_cols$cluster][870:872,870:872]
```




